---
title: "Data Frame Exploration"
author: "Emre Erhan"
date: "September 17, 2018"
output: "github_document"
---

## Exploring `gapminder`

### Setting up

We load `gapminder` using the `library` function

```{r}
library(gapminder)
```

### Exploring gapminder

The `head` command gives us the first part of the gapminder dataset, so we can see what it looks like.
```{r}
head(gapminder)
```
We can get a good summary of `gapminder` using the summary function:
```{r}
summary(gapminder)
```

### Let's look at population sizes

Let's look at some descriptive statistics for the `pop` column.
```{r}
median(gapminder$pop)
mean(gapminder$pop)
sd(gapminder$pop)
var(gapminder$pop)
```
We can see that these statistics agree with the statistics we got from the summary command above.

### An `if` statement

Let's see if the mean life expectancy is greater than that of Canada's, 82.
```{r}
if (mean(gapminder$lifeExp) > 82){
  print('It is!')
}else{
  print('It\'s not!')
}
```


### Let's see how Canadian life expectancy has changed over the years

```{r}
canada_index = gapminder$country=="Canada"
x = gapminder$year[canada_index]
y = gapminder$lifeExp[canada_index]
scatter.smooth(x, y, xlab="Years", ylab="Life expectancy")
```

### Is there a relationship between the `gdpPercap` and `pop` of Turkey over the years the data was taken?

```{r}
turkey_index = gapminder$country=="Turkey"
x = gapminder$pop[turkey_index]
y = gapminder$gdpPercap[turkey_index]
scatter.smooth(x, y, xlab="Population of Turkey", ylab="GDP/capita of Turkey")
```

There does appear to be a relationship! We can confirm this by getting the correlation.
```{r}
cor(x, y)
```

However, remember that a correlation does not suggest that either of these variables impact each other directly. We cannot say anything causal.