---
title: "Data Frame Exploration"
output: 
    html_document:
        theme: cerulean
        toc: true
        keep_md: true
---

## Exploring `gapminder`

### Setting up

We load `gapminder` using the `library` function

```{r}
library(gapminder)
```

### Exploring gapminder

The `head` command gives us the first part of the gapminder dataset, so we can see what it looks like.
```{r}
head(gapminder)
```
We can get a good summary of `gapminder` using the summary function:
```{r}
summary(gapminder)
```

### Let's focus on the population sizes
```{r}

```
Let's look at some descriptive statistics for this column.
```{r}
median(gapminder$pop)
mean(gapminder$pop)
sd(gapminder$pop)
var(gapminder$pop)
```
We can see that these statistics agree with the statistics we got from the summary command above.

### An `if` statement

Let's see if the mean life expectancy is greater than that of Canada's, 82.
```{r}
if (mean(gapminder$lifeExp) > 82){
  print('It is!')
}else{
  print('It\'s not!')
}
```
